Sustainable Pathways for Irish homes in changing TIMES

### 2021-06-28 manuscript.tex
Done -

In 2014, Rogan et al. \cite{Rogan2014LEAPsSystem} used OSeMOYOS to explore residential energy service demands in Ireland
 LEAP writes an input file for OSeMOSYS, runs it and imports its results back into LEAP. Currently, the LEAP input file only contains the aggregated electricity demand as the main driver for the optimization within OSeMOSYS.
 the model examines the
 3 main sceanrios - exploring transtport, residential, services and industy. The reference sceanrio - 
REF - TFC increase by 6.7\% in 2020, base year is 2008. space and water heating TFC have decreased in share of total residential TFC from 82 to 77 \%, whereas lighting, cooking and appliances have increased in TFC from 18 to 23 \%. assume that 1.4 \% of existing dwellings per annum will undergo a shallow retrofit in the reference
case, corresponding to just below 300,000 dwellings

In the energy efficiency scenario, we assume that 800,000 residential dwellings will be retrofitted between 2009 and 2020, in line with government targets for the BEH scheme,  Other policies in the residential sector are building regulations 2011 and building regulations 2013 and CFL lighting,  The combined impact of these policies is a reduction in TFC of 9.9 \% compared to the reference scenario.

The energy efficiency+scenario explores the impact of an ambitious retrofitting scenario beyond its current rate of implementation: retrofitbest is realised by retrofitting the same overall number of dwellings but undertaking a deeper, more comprehensive package of
energy efficiency measures on each, resulting in 7.2 \% savings compared to the energy efficiency scenario and 16.5 \% savings compared to the reference scenario.
  Insights - A challenging issue for energy modelling of policy is the
barriers and obstacles to investment, particularly in
residential energy efficiency measures such as split
incentives and information gaps, which lead to suboptimal
uptake of economically beneficial energy
efficiency improvements.

Thellufsen examining DH with EnergyPLAN and TIMES in 2019 \cite{Thellufsen2019ImplementingIreland}.

In Ireland few models have used energy system optimisation model to explore the residential sector. The majority of previous modelling on the residential sector in Ireland has involved bottom-up residential sector only models, rather than energy system models specifically examining the residential sector. Nevertheless, the residential models have provided insights on input data and methodology approach used in Ireland, some of which are outlined below.

Early work by Healy and CLinch ---\cite{Clinch2000HousingMortality,Healy2002FuelIreland,Clinch2003ValuingModel,Clinch2000DomesticFailure,Clinch2001Cost-benefitEfficiency}

@@

SEAI’s National Energy Modelling Framework (NEMF) is a suite of modelling tools used for modelling the Irish
energy system, both supply and demand side, to inform Irish Government energy policy and project Ireland’s
energy use based on these policie - https://www.seai.ie/publications/National-Energy-Projections-to-2030.pdf
The macro-economic model currently in use is called COSMO (Core Structural Model of the Irish Economy),
which creates a baseline forecast of energy consumption by sector (ESRI, 2017c). A range of sectoral models
are used to analyse both supply and demand side energy. Electricity supply is modelled using a commercial
power system software called PLEXOS47. A BioHeat model is used for techno-economic modelling of
bioenergy supply chains in Ireland for the electricity generation, transport and heat sectors48. There is also a
separate transport energy simulation model, which considers biofuels blending and the growth of low
emissions vehicles in the vehicle stock, including electric vehicles. In addition, separate modelling by an
energy-efficiency policy analysis tool (Unlocking model) informed the energy efficiency estimates included in
the NEMF49.

@@

\subsection{Limitations}
\label{Dis:Limit}
Consumers’ final investment decision can be
represented by “willingness-to-pay” curves – for
instance, around 10\% of owner-occupiers are willing
to invest in energy efficiency measures for a payback
time of 4 years, falling to 0\% for a payback time of
6 years for some consumer cohorts
Type of Tenure is ignored
BER aggeragateed
housing stock u-values
behaviour and occupancy
Area
expected energy consumption weakness of BER database as 
Actual energy consumption for dwellings of
similar or near identical construction may vary by up to 600%
depending on usage \cite{Sunikka-Blank2012BuildingConsumption}
\section{Conclusions}
\label{Conclusion}
This paper reports results on ambitious mitigation target in the period to 2050 for the Irish energy system. The analysis have been performed using the Irish TIMES model
\textit{We stress the importance of differentiating between technical con-
straints and behavioural constraints in optimisation models. Including both conflates what is considered desirable under the economic welfare paradigm (i.e. cost-minimising decisions) with “realistic decisions”, complicating the interpretation of the results. Optimisation models are by design well suited to study cost-optimal behaviour, and running scenarios just with technical constraints facilitates the identification of areas where market barriers are more entrenched, and interventions to accelerate the uptake of low-carbon practices may be needed.} \cite{Astudillo2017CanModelb}
\section{Acknowledgements}
\label{sec:Acknowledgements}
Tomas %CAPACITY is funded under the Climate Action Modelling Group (CAMG) by DECC
% https://www.marei.ie/project/capacity/
CHIMERA is supported by a research grant from Science Foundation Ireland (SFI) and the National Natural Science Foundation of China (NSFC) under the SFI-NSFC Partnership Programme, grant no. 17/NSFC/5181 and
supported by MaREI, the SFI Research Centre for Energy, Climate, and Marine [Grant No: 12/RC/2302\_P2]

>>>>

In 2014, Rogan et al. \cite{Rogan2014LEAPsSystem} linked OSeMOYOS, an ESOM to the Long Range Energy Alternatives Planning System (LEAP) simulation model, to explore specific energy efficiency policies not only in the residential sector, but also in transport, industry and services sector in Ireland. The bottom up development of the residential sector in LEAP was combined with the cost optimisation capabilities of OSeMOYOS. Three scenarios are explored, namely, the reference scenario, energy efficiency scenario and energy efficiency+ scenario. The reference assume that 300,000 existing dwellings will undergo a shallow retrofit from 2009 to 2020. In the reference scenario the residential total energy consumption rose 6.7 \% in 2020, compared to 2008. The share of space and water heating decreased from 82 \% to 77 \% in that period. In the energy efficiency scenario 800,000 residential dwellings are retrofitted between 2009 and 2020, combined with the effect of other policies such as building regulations and lighting, resulted in a reduction in of energy consumption of 9.9 \% in 2020, compared to the reference scenario. The energy efficiency+ scenario also allows for 800,000 residential dwellings to be retrofitting, but accounts for deeper retrofits. The results in a reduction in of energy consumption of 16.5 \% in 2020, compared to the reference scenario.Rogan et al. discussed the barriers, obstacles to investment, split incentives and information gaps which restrict energy efficiency improvements in the residential sector. The importance for policy to not only look at number of retrofit, but also the depth of retrofit works, is shown by the difference in scenarios. The model also found considerable scope for energy savings to be made through improved building regulations.

In 2018, Thellufsen et al. \cite{Thellufsen2019ImplementingIreland} investigated the potential for the implementation of district heating in both the commercial and residential sectors of Ireland. The paper analyses the $CO_2-80$ scenario ( 80 \% $CO_2$ reduction in 2050, compared to 1990 ) with and without district heating using the previous Irish TIMES model and EnergyPLAN. The previous version of Irish TIMES did not account for district heating, so EnergyPLAN is used to investigate the operation and  feasibility of district heating in Ireland. A sensitivity analysis was preformed on investment costs, heat grid losses and discount rates, however in all cases the inclusion of district heating is optimal. The final results find district heating reduces fuel consumption by 3.5 \% and the conversion of 37 \% of the Irish heat demand to district heating is found to be optimal. Both the TIMES model and climate target are superseded and the study did not include the future heat savings expected from retrofitting, nevertheless this study has provided insights on the potential of district heating.

\par In Ireland few models have used energy system optimisation model to explore the residential sector. The majority of previous modelling on the residential sector in Ireland has involved bottom-up residential sector only models, rather than energy system models specifically examining the residential sector. Nevertheless, the residential models have provided insights on input data and methodology approach used in Ireland, some of which are outlined below.

Clinch and Healy analysed different aspects of thermal comfort in the Irish residential sector. This was the first comprehensive analysis of the residential sector, their studies examined fuel poverty \cite{Healy2002FuelIreland}, excess winter mortality \cite{Clinch2000HousingMortality}, market failures \cite{Clinch2000DomesticFailure} and energy efficiency  \cite{Clinch2003ValuingModel,Clinch2000DomesticFailure,Clinch2001Cost-benefitEfficiency,Clinch2001ModellingEfficiency}. In 2000 Clinch and Healy developed and published a domestic energy efficiency model to analyse the cost-benefit of energy efficiency \cite{Clinch2001Cost-benefitEfficiency}.The Energy-Assessment Model (EAM) is a bottom-up model developed in Mircsoft Excel which incorporates 1,824 representative dwellings ( 8 building archetypes, 12 levels of insulation and 19 heating systems ), this was not linked to other sectors of the economy and only looked at space and water heating. A cost-benefit model (CBM) was fitted to the EAM to facilitate the conversion of the physical estimates into monetary amounts. Under the predicted scenario, energy savings easily outweighing programme costs, representing a benefit-cost ratio of 1.7.The benefits of reduced sickness, hospitalisation and drugs and mortality, were quantified, this moved the  benefit-cost ratio up to 3.0.
The same model was used by Clinch, Healy and King \cite{Clinch2001ModellingEfficiency} in addition three models of householder spending on energy are developed here to to evaluate the human behavioural component
into the computer model. the modelling process involved developing a computer model of the energy performance of the existing dwelling stock. It was noted, energy consumption predicted by the model is considerably higher than given in the national energy balance, due to intertal temperature assumptions. 
The model showed the energy-efficiency programme ( all houses retrofitted to 1997 thermal building standards in 10 years ) represents a saving of 24 \% when compared to the reference scenarios.

@@

Finnaly, it is worth mentioning SEAI’s National Energy Modelling Framework (NEMF), which is a suite of modelling tools used for modelling the Irish energy system. Some of the main models used in the NEMF include:
\begin{itemize}
\item {Core Structural Model of the Irish Economy (COSMO)  is a macro-economic model used by the Economic and Social Research Institute (ESRI) to create a baseline forecast of energy consumption by sector} \cite{Bergin2017COSMO:Ireland}
\item{PLEXOS is a power systems modeling tool used for electricity modeling and planning}\cite{PLEXOSExemplar}
\item{Energy-efficiency policy analysis tool informs the energy efficiency estimates} \cite{Scheer2015UnlockingTob}
\begin{itemize}
\item{Consumer Choice Model determines the attractiveness of energy efficiency finance options}\cite{SustainableEnergyAuthorityofIreland2017BehaviouralSector}
\end{itemize}
\item{BioHEAT is a techno-economic simulation model of bioenergy and heat, used by SEAI to support policy decision in Ireland.}\cite{Durusut2018BioHEAT:Sectors}
\end{itemize}

The listed NEMF models and more are used by SEAI to model the impact of incentives and opportunities created by specific policy instruments. The models are a combination of top down, ..........

@@

\subsection{Limitations}
\label{Dis:Limit}
%Discrepancies were investigated, and assumptions reviewed intuitively in order to achieve a good match.=
Consumers’ final investment decision can be
represented by “willingness-to-pay” curves – for
instance, around 10\% of owner-occupiers are willing
to invest in energy efficiency measures for a payback
time of 4 years, falling to 0\% for a payback time of
6 years for some consumer cohorts
Type of Tenure is ignored
BER aggeragateed
housing stock u-values
behaviour and occupancy
Area
expected energy consumption weakness of BER database as 
Actual energy consumption for dwellings of
similar or near identical construction may vary by up to 600%
depending on usage \cite{Sunikka-Blank2012BuildingConsumption}
\section{Conclusions}
\label{Conclusion}
This paper reports results on ambitious mitigation target in the period to 2050 for the Irish energy system. The analysis have been performed using the Irish TIMES model
\textit{We stress the importance of differentiating between technical con-
straints and behavioural constraints in optimisation models. Including both conflates what is considered desirable under the economic welfare paradigm (i.e. cost-minimising decisions) with “realistic decisions”, complicating the interpretation of the results. Optimisation models are by design well suited to study cost-optimal behaviour, and running scenarios just with technical constraints facilitates the identification of areas where market barriers are more entrenched, and interventions to accelerate the uptake of low-carbon practices may be needed.} \cite{Astudillo2017CanModelb}
\section{Acknowledgements}
\label{sec:Acknowledgements}
Tomas %CAPACITY is funded under the Climate Action Modelling Group (CAMG) by DECC
% https://www.marei.ie/project/capacity/
CHIMERA is supported by a research grant from Science Foundation Ireland (SFI) and the National Natural Science Foundation of China (NSFC) under the SFI-NSFC Partnership Programme, grant no. 17/NSFC/5181 and
supported by MaREI, the SFI Research Centre for Energy, Climate, and Marine [Grant No: 12/RC/2302\_P2]
